---
title: "Tu-Linear-02"
author: "Sicheng Peng, Hajar El Mesaoudi"
date: "2023/7/4"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,results = "hide",warning=FALSE}
library("readxl")
linear_data = read_excel("./Tutorial/1_LinearRegression_Tutorial_Data.xlsx")
```
We squared Ctenure and Mtenure and used both as DVs. The result shows both them are significant drivers for financial profit.
```{r}
Ctenure2 = linear_data$Ctenure^2
Mtenure2 = linear_data$Mtenure^2
performancelmsq = lm(Profit~Ctenure+Ctenure2+Mtenure+Mtenure2+Pop+Comp+Visible+PedCount+Res+Hours24+CrewSkill+MgrSkill+ServiceQuality, data = linear_data)
summary(performancelmsq)
```
Based on the previous results, we find Visible may not be a relative DV for profit. We removed Visible from the linear regression function and the results had no major changes. Hence we conclude Visible is not a significant driver for profit.
```{r}
performance = lm(Profit~Ctenure+Ctenure2+Mtenure+Mtenure2+Pop+Comp+PedCount+Res+Hours24+CrewSkill+MgrSkill+ServiceQuality, data = linear_data)
summary(performance)
```
