---
title: "Tu-Linear-01"
author: "Sicheng Peng, Hajar El Mesaoudi"
date: "2023/7/4"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
In order not to be bothered with rounding the numbers, set `options(digits=3)`
`r options(digits=3)`

## Exercise 01
Install package dependencies and read dataset for tutorial 1 as linear_data
```{r,results = "hide",warning=FALSE}
#install.packages("readxl")
library("readxl")
linear_data = read_excel("./Tutorial/1_LinearRegression_Tutorial_Data.xlsx")
```
We perform linear regression on profit, using crew tenure and manager tenure as DVs.
The result p-value for Mtenure is lower than 0.05, thus H0 is accepted, manager tenure is a driver of store level financial performance. The result p-value for Ctenure is higher than 0.1, thus H0 is rejected, customer tenure is not a significant driver of store level financial performance. 
```{r}
performancelm = lm(linear_data$Profit~linear_data$Mtenure+linear_data$Ctenure)
summary(performancelm)
```
## Exercise 02
We try different DVs.
Result shows DV Visible and CrewSkill are not significant in related to financial performance, ServiceQuality is partially relative. The rest DVs are all significant drivers of financial performance.
```{r}
totalperformancelm = lm(Profit~Mtenure+Ctenure+Pop+Comp+Visible+PedCount+Res+Hours24+CrewSkill+MgrSkill+ServiceQuality, data = linear_data)

summary(totalperformancelm)
```



